LME-MPM applied to quasi-brittle fracture.

### 2020-06-13 2020_EFM_MPM_Eigensoftening.tex
Cambios MMP propuestos por DM

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{efgm}{EFGM}{Element-Free Galerkin Method}
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}
\newacronym{npc}{NPC}{Newmark Predictor-Corrector}
\newacronym{fe}{FE}{Forward Euler}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
The presence of cracks is a violation of the continuity requirement of the
\acrfull{fem}. To overcome this issue, numerous numerical
artifacts have been proposed with the aim of reproducing such a
complex behaviour. These techniques vary from the employment of cohesive
approaches~\cite{Barenblatt,Hilleborg_1976}, \PNA{to the gradual insertion of}
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries, or handling arbitrary crack paths by level set
representation of the fracture surface~\cite{Belytschko_03}. The
simulation of fracture propagation in a more accurate and effective
way can be considered as one of the \PNA{ongoing goals} of the development of novel spatial discretization methods like meshfree techniques. Some
examples of it are the \acrfull{mpm}
\cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the
\acrfull{efgm} \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013}, the
\acrfull{sph} \cite{Wang_2020,Wang_2019}, the \acrfull{otm} \cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or Peridynamics
\cite{HA_2011,RABCZUK_2017} among others. \\

Regarding \acrshort{mpm}, fracture \PNA{has been treated numerically through two different ways}. On the one hand, the ``CRAcks with Material Points (CRAMP)'' is proposed~\cite{Nairn_2003,Nairn_2006}. \PNA{The basis of this methodology consist of the removal of} the restriction
of the single-valued velocity field close to the crack through, at least, two sets of nodes. In this method, different labels are assigned to the material points
and nodes to distinguish if they are in the same side of the crack or
not. Under this approach, crack surface is described with line segments
in 2D and triangle patches in 3D cases. The chosen criteria for crack
propagation is based on such parameters as energy release rate
analyzed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the stress
intensity factor or the J-integral discussed by Guo \& Nairn
(2004)\cite{Nairn_2004}. 
On the other hand, the \PNA{foundation} of the second approach is the introduction of failed
material points to describe the crack evolution. In these methods, the
formation of failed points describes the nucleation of cracks, and
thereafter its propagation and branching. Consequently, the position
of the crack does not need to be explicitly stated. This fact represents
significant advantages over the ``CRAMP''. Under this approach, the
prediction of failure evolution is computed with a decohesion model,
which has been discussed by Chen {\it et al.}\cite{Zhenmao_2005} and
Schreyer {\it et al.}\cite{Schreyer_2002}. Successful simulations are provided within the framework of the fracture of brittle materials by Chen {\it et
  al.} \cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}.\\

@@

First, the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. \PNA{Although the traditional \acrshort{mpm} implementation allows collision of two bodies, inaccurate energy conservation is obtained without the implementation of a conservative contact algorithm to reproduce the hammer impact in the beam. %In the
%%\acrshort{mpm}, the implementation of this feature can be bypassed to obtain preliminary results. Nevertheless, this kind of algorithm could
%be implemented to get ever more accurate solutions. 
Despite the observed discrepancies, this is
out of the scope of this document, although future research of authors will overcome this
limitation. Similarly, results of support forces will be more realistic when modelled by a suitable contact algorithm instead of avoiding any vertical displacement}.\\

@@

\section{Conclusions}
\label{sec:4}
\PNA{Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.}
\cite{Zhang_EE_2020}, two additional improvements have been proposed in the present research in order to 
solve the shortcomings of the original concept that Zhang {\it et al.} pointed out in the aforementioned paper \cite{Zhang_EE_2020}.

Firstly, by
introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrfull{fem}. Furthermore, this failure criterion admits the design of more elaborated damage curves, which can be employed in order to describe the failure process in more sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced concretes\cite{Ruiz2019}. Interesting researches have been done on this field~\cite{Navas_2018_ES}. In addition, the eigensoftening approach can be extrapolated to rock fracture in the same manner the eigenerosion was~\cite{Wang2017}, taking into account that the volumetric stiffness remains in order to support bulk loads.

Secondly, \acrshort{lme} is employed to remove from
the solution those wiggles provoked by inaccuracies in the stress integration. This spatial discretization was previously employed within the \acrshort{mpm} framework with excellent results under the elastic regime~\cite{Wobbes2020,Molinos2020}. The traditional time scheme of the \acrshort{mpm} has been improved with a Newmark Predictor Corrector scheme as well. In this research, the aforementioned techniques have helped to improve the
results in fracture propagation through the eigensoftening and eigenerosion algorithms. Further improvement may be reached by adapting \acrshort{lme} through the
deformation gradient \cite{Kochmann2019}. \textit{A priori} this would enhance the localisation properties of the proposed algorithms in more sophisticated domains.

Finally, for better capturing the measurement of forces made in the
laboratory, more realistic contact tools have to be
developed. Navas~\textif{et al.} demonstrate the importance of the
stiffness of the support in order to obtain accurate results of the
reaction forces. Interesting contact algorithms within the
\acrshort{mpm} methodology are available in the literature
\cite{Bardenhagen_Contact_2001,XZhang_Contact_2011}}

>>>>

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{xfem}{XFEM}{Extended Finite Element Method}
\newacronym{efgm}{EFGM}{Element-Free Galerkin Method}
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}
\newacronym{npc}{NPC}{Newmark Predictor-Corrector}
\newacronym{fe}{FE}{Forward Euler}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
\DM{The failure mechanisms of quasi-brittle materials and the  crack propagation have been studied from different fields such as materials science, structural engineering, earth science and petroleum engineering . The prediction of crack propagation has been approached with different numerical techniques. The \acrfull{fem} is widely used in industrial application. However,} the presence of cracks is a violation of the continuity requirement of the \acrshort{fem}. To overcome this issue, numerous numerical
artifacts have been proposed with the aim of reproducing such a
complex behaviour. These techniques vary from the employment of cohesive
approaches~\cite{Barenblatt,Hilleborg_1976}, or \PNA{to the gradual insertion of}
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries. To alleviate these shortcomings of the
\acrshort{fem}, the \acrfull{xfem} in conjunction with level set
technique was introduced by Belytschko {\it et al.}
\cite{Belytschko_03} to handle arbitrary crack paths level
set. \MMP{Some relevant contributions to the \acrshort{xfem} are the
modeling micro-scale crack propagation \cite{HUANG_2018}, the
introduction of a cohesive zone model \cite{NAGASHIMA_2016}, and more
recently improving it capability to predict the crack path in 3D cases
\cite{YUN_2019}.}

The simulation of fracture propagation in a more accurate and effective
way can be considered as one of the \PNA{ongoing goals} of the development of novel spatial discretization methods like meshfree techniques. Some
examples of it are the \acrfull{mpm}
\cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the
\acrfull{efgm} \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013}, the
\acrfull{sph} \cite{Wang_2020,Wang_2019}, the \acrfull{otm} \cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or Peridynamics
\cite{HA_2011,RABCZUK_2017} among others. \\

\MMP{Concerning} \acrshort{mpm}, fracture \PNA{has been treated numerically through two different ways}. On the one hand, the ``CRAcks with Material Points (CRAMP)'' is proposed~\cite{Nairn_2003,Nairn_2006}. \PNA{The basis of this methodology consist of the removal of} the restriction
of the single-valued velocity field close to the crack through, at least, two sets of nodes. In this method, different labels are assigned to the material points
and nodes to distinguish if they are in the same side of the crack or
not. Under this approach, crack surface is described with line segments
in 2D and triangle patches in 3D cases. The chosen criteria for crack
propagation is based on such parameters as energy release rate
analyzed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the stress
intensity factor or the J-integral discussed by Guo \& Nairn
(2004)\cite{Nairn_2004}. 
On the other hand, the \PNA{foundation} of the second approach is the introduction of failed
material points to describe the crack evolution. In these methods, the
formation of failed points describes the nucleation of cracks, and
thereafter its propagation and branching. Consequently, the position
of the crack does not need to be explicitly stated. This fact represents
significant advantages over the ``CRAMP''. Under this approach, the
prediction of failure evolution is computed with a decohesion model,
which has been discussed by Chen {\it et al.}\cite{Zhenmao_2005} and
Schreyer {\it et al.}\cite{Schreyer_2002}. Successful simulations are provided within the framework of the fracture of brittle materials by Chen {\it et
  al.} \cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}.\\

@@

First, the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. \PNA{Although the traditional \acrshort{mpm} implementation allows collision of two bodies, inaccurate energy conservation is obtained without the implementation of a conservative contact algorithm to reproduce the hammer impact in the beam.
Despite the observed discrepancies, this is out of the scope of this
document, although future research of authors will overcome this
limitation. Similarly, results of support forces will be more
realistic when modelled by a suitable contact algorithm instead of
avoiding any vertical displacement}.\MMP{Navas {\it et al.}
(2017)\cite{Navas_2017_ES} demonstrate the importance of the stiffness
of the support in order to obtain accurate results of the reaction
forces. Interesting contact algorithms within the \acrshort{mpm}
methodology are available in the literature
\cite{Bardenhagen_Contact_2001,XZhang_Contact_2011}.}

@@

\section{Conclusions}
\label{sec:4}

\DM{The paper proposes the eigensoftening algorithm in a \acrshort{MPM} framework for quasi-brittle materials to overcome the limitations observed in the EigenMPM \cite{Zhang_EE_2020}. The overestimation of tensile 
stress and strain peaks in the fracture mechanism of quasi-brittle
materials and the stress oscillations observed with conventional
\acrshort{mpm}/\acrshort{gimp} are faced.} \PNA{Based on the
EigenMPM approach to fracture, two additional improvements have been proposed in the present research in order to solve the shortcomings of the original concept pointed out in the aforementioned paper \cite{Zhang_EE_2020}.}

Firstly, by introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrshort{fem}. Furthermore, this failure
criterion admits the design of more elaborated damage curves, which
can be employed in order to describe the failure process in more
sophisticated materials. Secondly to achieve better results, three new
numerical algorithms has been implemented. The \acrshort{lme}
approximants are employed to remove from
the solution those wiggles provoked by inaccuracies in the stress
integration. The traditional time scheme of the
\acrshort{mpm} has been improved with a Newmark Predictor Corrector
scheme as well. And the traditional concept of element connectivity
has been removed to allow a more flexible and efficient implementation of search
algorithm. In this research, the aforementioned techniques have helped
to improve the results in fracture propagation through the
eigensoftening and eigenerosion algorithms.

Finally, further improvements of the present research are exposed. The
algorithm would have the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced
concretes\cite{Navas_2018_ES,Ruiz2019}. In addition, the eigensoftening
approach can be extrapolated to rock fracture in the same manner the
eigenerosion was~\cite{Wang2017}, taking into account that the
volumetric stiffness remains in order to support bulk loads. Also the
benefits of \acrshort{lme} can be exploited by adapting it through the
deformation gradient \cite{Kochmann2019}. \textit{A priori} this would
enhance the localisation properties of the proposed algorithms in more
sophisticated domains.

### 2020-06-13 _region_.tex
Cambios MMP propuestos por DM

\section{Conclusions}
\label{sec:4}
\PNA{Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.}
\cite{Zhang_EE_2020}, two additional improvements have been proposed in the present research in order to 
solve the shortcomings of the original concept that Zhang {\it et al.} pointed out in the aforementioned paper \cite{Zhang_EE_2020}.

Firstly, by
introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrfull{fem}. Furthermore, this failure criterion admits the design of more elaborated damage curves, which can be employed in order to describe the failure process in more sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced concretes\cite{Ruiz2019}. Interesting researches have been done on this field~\cite{Navas_2018_ES}. In addition, the eigensoftening approach can be extrapolated to rock fracture in the same manner the eigenerosion was~\cite{Wang2017}, taking into account that the volumetric stiffness remains in order to support bulk loads.

@@

\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1280) }

>>>>

\section{Conclusions}
\label{sec:4}

\DM{Cometario general: Pienso que hay que estructurar un poco la conclusión. No agregar tantas citas. Acá estás concluyendo tu trabajo. Iría de lo general a lo particular. Pondría claramente en cada parrafo cual es el aporte. Con tantas citas no queda claro que hiciste vos y que los otros. Las citas estarán puestas cuando vos comenzaste a hablar de lo que hiciete en el cuerpo del documento. Acá las citas minimas y tus aportes.
Imagino un primer parrafo asi diciendo cuales son las limitaciones ...propuesta modificable...:}

\DM{The paper proposes the eigensoftening algorithm in a \acrshort{MPM} framework for quasi-brittle materials to overcome the limitations observed in the EigenMPM \cite{Zhang_EE_2020}. The overestimation of tensile 
stress and strain peaks in the fracture mechanism of quasi-brittle materials and the stress oscillations observed with conventional \acrshort{mpm}/\acrshort{gimp} are faced.  }

\PNA{Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.} 
\cite{Zhang_EE_2020}, two additional improvements have been proposed in the present research in order to 
solve the shortcomings of the original concept that Zhang {\it et al.} pointed out in the aforementioned paper \cite{Zhang_EE_2020}. }

Firstly, by
introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrshort{fem}. Furthermore, this failure criterion admits the design of more elaborated damage curves, which can be employed in order to describe the failure process in more sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced concretes\cite{Ruiz2019}. Interesting researches have been done on this field~\cite{Navas_2018_ES}. In addition, the eigensoftening approach can be extrapolated to rock fracture in the same manner the eigenerosion was~\cite{Wang2017}, taking into account that the volumetric stiffness remains in order to support bulk loads.

@@

\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1295) }
